---
title: "Thetas over time"
output:
    html_document:
    df_print: paged
theme: sandstone
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(viridis)
library(broom)
library(hrbrthemes)
theme_set(theme_ipsum())

knitr::opts_chunk$set(tidy = FALSE,
                      fig.width = 6,
                      fig.height = 5)

# library(thematic)
```

## Thetas

```{r read}
# ws_t_raw = read_csv(here::here("theta-csvs/theta_table_whosampled.csv.gz"))
degrees = read_csv(here::here("graph-csvs/ws-degrees.csv"))

# TEMPORARY:
wsp = read_csv(here::here("centrality-csvs/whosampled_pagerank.csv"))
wsd = read_csv(here::here("centrality-csvs/whosampled_disruption.csv"), 
               col_types = cols(
                 .default = col_double(),
                 X1 = col_character()
               ))

wsp = wsp %>% 
  rename(artist = X1) %>% 
  pivot_longer(-artist, names_to = "ano", values_to = "value") %>% 
  mutate(name = artist) %>% 
  group_by(ano) %>% 
  mutate(theta = rank(value, ties.method = "average")/ n(), metric = "pagerank") %>% 
  ungroup() %>% 
  select(-value)

wsd = wsd %>% 
  rename(artist = X1) %>% 
  pivot_longer(-artist, names_to = "ano", values_to = "value") %>% 
  filter(!is.na(value)) %>% 
  mutate(name = artist) %>% 
  group_by(ano) %>% 
  mutate(theta = rank(value, ties.method = "average") / n(), metric = "disruption") %>% 
  ungroup() 

glimpse(wsd)

ws_t_raw = bind_rows(wsp, wsd)
ws_t_raw = ws_t_raw %>% 
  mutate(ano = as.numeric(ano))
rm(wsd, wsp)
```


```{r}
glimpse(ws_t_raw)
```

```{r}
#skimr::skim(ws_t_raw)
```

### Transform

```{r}
end_ranks = ws_t_raw %>% 
  filter(ano == max(ano)) %>% 
  group_by(metric) %>% 
  select(artist, metric, end_theta_i = theta) %>% 
  ungroup()

ws_t = ws_t_raw %>% 
  filter(ano >= 1960) %>% 
  left_join(end_ranks, by = c("metric", "artist")) %>% 
  mutate(name = artist) %>% 
  group_by(metric, ano) %>% 
  ungroup() %>% 
  left_join(degrees, by = c("artist", "ano" = "year")) %>% 
  filter(out_degree >= 5, in_degree >= 5)

ws_t_wide = ws_t %>% 
  select(artist, name, ano, theta, metric) %>% 
  pivot_wider(names_from = metric, values_from = theta)

with_history = ws_t %>% 
  count(metric, artist) %>%
  filter(n >= 5) %>%    # <-- DIFFERENT FROM ALLMUSIC
  distinct(metric, artist) %>% 
  count(artist) %>% 
  filter(n == 2) %>% 
  pull(artist)

ws_t_h = ws_t %>% 
  filter(artist %in% with_history)

ws_t_h_wide = ws_t_wide %>% 
  filter(artist %in% with_history)
```

### Explore

Some exemplary artists

```{r}
few_artists = function(d, n){
  d %>% 
    filter(metric == "disruption") %>% 
    sample_n(n) %>% 
    pull(artist)
}

examples_high_t = ws_t_h %>% 
  group_by(metric, artist, name) %>%
  summarise(avg = mean(theta), count = n()) %>%
  filter(count > 2) %>%
  ungroup %>% 
  group_by(metric) %>% 
  top_n(3, avg) 
examples_high_t

examples_sd_t = ws_t_h %>% 
  group_by(metric, artist, name) %>%
  summarise(sd = sd(theta)) %>%
  ungroup %>% 
  group_by(metric) %>% 
  top_n(3, sd) 


model_t = function(d){
  lm(theta ~ ano, data = d)
}

# models = ws_t_h %>%
#   group_by(artist, metric) %>% 
#   nest() %>% 
#   mutate(model = map(data, model_t))
# 
# coefs = models %>% 
#   mutate(coefs = map(model, tidy)) %>% 
#   unnest(coefs)  %>% 
#   filter(term == "ano")
# 
# examples_grow_disrupt = coefs %>% 
#   select(artist, metric, estimate) %>% 
#   filter(metric == "disruption") %>% 
#   ungroup %>% 
#   top_n(4, estimate)
# 
# examples_grow_disrupt
```

Everything over time

```{r}
plot_traj_sbs = function(d, artists = c(), names = c()){
  d %>% 
    filter(artist %in% artists | name %in% names) %>%
    ggplot(aes(x = ano, y = theta, color = name, group = artist)) +
    facet_wrap(~ metric) + 
    geom_line() + 
    labs(x = "year", y = expression(1-theta))
}

set.seed(1)

ws_t %>% 
  plot_traj_sbs(few_artists(., 5))

# ws_t %>% 
#   plot_traj_sbs(examples_grow_disrupt$artist)

ws_t %>% 
  plot_traj_sbs(names = c("Kraftwerk", "John Cage", "Sun Ra"))

ws_t %>% 
  plot_traj_sbs(names = examples_sd_t$name) 
```

All thetas over time together

```{r}
# ws_t %>% 
#   filter(metric == "pagerank") %>% 
#   ggplot(aes(x = ano, y = end_theta_i, fill = theta)) + 
#   facet_grid(. ~ metric) +
#   geom_tile() + 
#   scale_fill_viridis(direction =  -1, option = "inferno")
# 
# ws_t %>% 
#   filter(metric == "disruption") %>% 
#   ggplot(aes(x = ano, y = end_theta_i, fill = theta)) + 
#   facet_grid(. ~ metric) +
#   geom_tile() + 
#   scale_fill_viridis(direction =  -1, option = "inferno")
```

#### Trajectories on a xy plane.

```{r}
ws_t_wide %>% 
  # filter(artist %in% few_artists(., 3)) %>%
  # filter(name %in% c("Kraftwerk")) %>%
  # filter(artist %in% examples_grow_disrupt$artist) %>%
  filter(name %in% examples_sd_t$name) %>%
  ggplot(aes(x = pagerank, y = disruption, color = name, group = artist)) +
  geom_line() + 
  geom_point(aes(size = ano, alpha = ano)) 
```

Are there correlations?

```{r}
ws_t_wide %>% 
  filter(!is.na(pagerank), !is.na(disruption)) %>% 
  ggplot(aes(x = pagerank, y = disruption)) +
  geom_hex()
  # geom_point(alpha = .2, size = .5) 
```

#### Some interesting trajectories

```{r}
ws_t_wide %>% 
  filter(name %in% c("Bukas Palad", "Caetano Veloso", "Shirley Bassey", "Kraftwerk")) %>% 
  ggplot(aes(x = pagerank, y = disruption, group = artist)) +
  facet_wrap(~ name) +
  geom_line() + 
  geom_point(aes(size = ano, alpha = ano)) + 
  scale_alpha_continuous(range = c(.2, .5))
```

---

## Surprises

```{r}
ws_s = read_csv(here::here("surprises-csvs/whosampled_surprises.csv.gz"))

ws_s = ws_s %>%
  mutate(
    surprise_disrupt = disrupt_surprise_hyp_growth_t +
      disrupt_surprise_hyp_past_t,
    surprise_pr = pagerank_surprise_hyp_growth_t + 
      pagerank_surprise_hyp_past_t,
    surprise_total = surprise_disrupt + surprise_pr
  )%>% 
  filter(surprise_total != 0)
```

```{r}
glimpse(ws_s)
```

Final dataframe

```{r}
whosampled =
  ws_t_h_wide %>%
  filter(!is.na(pagerank + disruption)) %>% 
  left_join(
    select(ws_s, name, year, matches("surprise")),
    by = c("ano" = "year", "name")
  ) %>% 
  filter(!is.na(surprise_total)) %>% 
  group_by(artist) %>% 
  arrange(ano) %>% 
  mutate(year_index = 0:(n()-1) / (n()-1)) %>% 
  ungroup()

glimpse(whosampled)
```

### Agregate views

Surprise x year index

```{r}
whosampled %>% 
  group_by(artist) %>% 
  mutate(i = 1:n()) %>% 
  pivot_longer(matches("surprise"), names_to = "surprise") %>% 
  ggplot(aes(x = i, y = value, group = i)) +
  facet_wrap(~surprise, ncol = 2) +
  # ggbeeswarm::geom_quasirandom(size = .3) +
  geom_boxplot(coef = 1000) +
  scale_y_log10(limits = c(1e-3, 1e3)) +
  labs(x = "Year index from appearance onwards")
```



```{r}
whosampled %>% 
  group_by(artist) %>% 
  mutate(i = 1:n()) %>% 
  ggplot(aes(x = i, y = surprise_total, group = i)) +
  geom_boxplot(outlier.color = NA) +
  scale_y_log10() + 
  labs(x = "Year index from appearance onwards")
```

```{r}
whosampled %>% 
  group_by(artist) %>% 
  arrange(ano) %>% 
  mutate(i = 1:n()) %>% 
  mutate(delta = surprise_total - lag(surprise_total)) %>%  
  ggplot(aes(x = i, y = delta, group = i)) +
  # ggbeeswarm::geom_quasirandom(size = .3) +
  geom_boxplot(coef = 1e9) + 
  # geom_boxplot(coef = 1000) +
  labs(x = "Year index from appearance onwards") + 
  ylim(-25, 25)
```


```{r}
whosampled %>% 
  ggplot(aes(x = surprise_pr, y = surprise_disrupt)) + 
  geom_point(alpha = .15, size = .25) + 
  scale_x_log10() + 
  scale_y_log10()

whosampled %>% 
  ggplot(aes(x = disrupt_surprise_hyp_past_t, y = disrupt_surprise_hyp_growth_t)) + 
  geom_point(alpha = .15, size = .25) + 
  scale_x_log10() + 
  scale_y_log10()


whosampled %>% 
  ggplot(aes(x = pagerank_surprise_hyp_past_t, y = pagerank_surprise_hyp_growth_t)) + 
  geom_point(alpha = .15, size = .25) + 
  scale_x_log10(limits = c(1e-6, 1e3)) + 
  scale_y_log10(limits = c(1e-6, 1e3))
```


#### highest surprise at some point


```{r}
plot_trajectory = function(artists, df){
  df %>% 
    filter(artist %in% artists) %>%
    mutate(first = year_index == 0) %>% 
    ggplot(aes(
      x = pagerank,
      y = disruption,
      color = year_index,
      fill = year_index,
      group = artist
    )) +
    facet_wrap( ~ name) +
    geom_line(size = 0.4, color = "black") +
    geom_point(aes(size = surprise_total), color = "black", alpha = .7, shape = 21) +
    scale_fill_viridis(option = "magma", direction = -1) +
    scale_size_area(trans = "sqrt", max_size = 10) + 
    labs(
      x = expression(1 - theta * " for PageRank"),
      y = expression(1 - theta * " for Disruption"),
      fill = "Prop. years\nactivity",
      size = "Surprise"
    )  + 
    guides(shape = F)
}
```


From all years, it's always the first.

```{r}
ws_t_h %>%
  filter(artist == "TIEKS")
```


```{r}
whosampled %>% 
  group_by(artist) %>% 
  summarise(surprise_total = max(surprise_total)) %>% 
  top_n(6, surprise_total) %>% 
  distinct(artist) %>% 
  pull(artist) %>% 
  plot_trajectory(whosampled)
```

```{r}
ws_t %>% 
  plot_traj_sbs(names = c("Pabllo Vittar"))
```

```{r}
whosampled %>% 
  filter(name == "Miles Davis") %>% 
  # filter(ano < 1995, ano > 1985) %>% 
  pivot_longer(c("pagerank", "disruption"), names_to = "metric", values_to = "value") %>% 
  ggplot(aes(x = ano, y = value, size = surprise_total)) + 
  facet_wrap(~ metric) +
  geom_point(alpha = .5) 
```

```{r}
whosampled %>% 
  filter(name == "Miles Davis") %>% 
  # filter(ano > 1985) %>% 
  pivot_longer(c("pagerank", "disruption"), names_to = "metric", values_to = "value") %>% 
  pivot_longer(matches("surprise"), names_to = "surprise", values_to = "suprise_value") %>% 
  ggplot(aes(x = ano, y = suprise_value)) + 
  facet_wrap(~ surprise, ncol = 2) +
  geom_point() + 
  geom_line()
```


```{r}
c("Taylor Swift") %>% 
  plot_trajectory(whosampled)
```


### Highest *after* first year


```{r}
whosampled %>% 
  group_by(artist) %>% 
  arrange(ano) %>% 
  mutate(delta = surprise_total - lag(surprise_total)) %>%  
  summarise(surprise_total = max(delta, na.rm = T)) %>% 
  top_n(9 , surprise_total) %>% 
  distinct(artist) %>% 
  pull(artist) %>% 
  plot_trajectory(whosampled) 
```

```{r}
whosampled %>% 
  filter(year_index == 1) %>% 
  top_n(100, wt = disruption) %>% View
```



```{r}
c("Taylor Swift",
  "Madonna",
  "Grandmaster Flash",
  "Massive Attack"
  # "Björk"
  # "DJ Shadow",
  # "Miles Davis"
  ) %>% 
  plot_trajectory(whosampled)

ggsave(here::here("ws-trajectories.pdf"), width = 5, height = 5)
```


#### highest average total surprise

```{r}
whosampled %>% 
  group_by(artist) %>% 
  summarise(surprise_total = mean(surprise_total)) %>% 
  top_n(6, surprise_total) %>% 
  distinct(artist) %>% 
  pull(artist) %>% 
  plot_trajectory(whosampled) +
  xlim(.3, 1)
```


Some interesting artists

```{r}
whosampled %>%
  filter(
    name %in% c(
      "Stevie Wonder",
      "Jorge Ben",
      "Shirley Bassey",
      "Fela Kuti",
      "Miles Davis"
    ),
    ano >= 1960
  ) %>%
  pull(artist) %>% 
  plot_trajectory(whosampled)  + 
  xlim(.3, 1)
```

```{r}
ws_s_wide = whosampled %>% 
  pivot_longer(cols = matches("surprise"), names_to = "surprise") 
  
# ws_s_wide %>% 
#   ggplot(aes(x = value)) + 
#   geom_density() +
#   facet_grid(surprise ~ ano) + 
#   scale_x_log10(limits = c(1e-4, 1)) 

whosampled %>% 
  ggplot(aes(x = surprise_pr, y = surprise_disrupt)) + 
  geom_point(alpha = .2, size = .5, color = "brown") + 
  facet_wrap(~ano) +
  scale_x_log10(limits = c(1e-6, 1e1)) + 
  scale_y_log10() +
  NULL
```

```{r}
ws_s_wide %>% 
  filter(str_starts(surprise, "surprise")) %>% 
  ggplot(aes(x = ano, y = value, group = ano, fill = surprise)) + 
  geom_violin() +
  geom_boxplot(width = .5, coef = 1e6, fill = "white") + 
  facet_grid(. ~ surprise) + 
  scale_y_log10(
                limits = c(1e-4, 1e4)
                ) +
  theme(legend.position = "None") +
  NULL
```

```{r}
ws_s_wide %>% 
  group_by(surprise, ano) %>% 
  summarise(value = sum(value)) %>% 
  ggplot(aes(x = ano, y = value, group = surprise)) + 
  geom_line() +
  geom_point() +
  facet_wrap(~ surprise, ncol = 2) 
```


```{r}
whosampled %>% 
  filter(ano >= 1960, metric == "pagerank") %>%
  ggplot(aes(x = surprise_pr, y = surprise_disrupt)) + 
  geom_point(alpha = .2, size = .5, color = "brown") + 
  facet_wrap(~ano) +
  scale_x_log10(limits = c(1e-6, 1e1)) + 
  scale_y_log10() +
  NULL
```

```{r}
rename_surprises = function(d){
  d %>% 
    mutate(
    s_class = if_else(grepl("disrupt",surprise), "Disruption", "Pagerank"),
    surprise = case_when(
      surprise == "pagerank_surprise_hyp_past_t" ~ "Past Rank",			
      surprise == "pagerank_surprise_hyp_growth_t"  ~ "Reg. Growth ",				
      surprise == "disrupt_surprise_hyp_past_t" ~ "Past Rank",			
      surprise == "disrupt_surprise_hyp_growth_t"~ "Reg. Growth ",				
      surprise == "surprise_disrupt" ~ "Disruption, sum",			
      surprise == "surprise_pr" ~ "Pagerank, sum"
    )
  )
}

ws_s_wide %>% 
  filter(grepl("surprise", surprise), !str_starts(surprise, "surprise")) %>% 
  rename_surprises() %>% 
  group_by(s_class, surprise, ano) %>% 
  summarise(value = sum(value)) %>% 
  ggplot(aes(x = ano, y = value, fill = surprise)) + 
  geom_area(alpha = .75) +
  facet_wrap(~ s_class, ncol = 1) + 
  theme(legend.position = "bottom",
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        ) +
  scale_fill_brewer(palette = "Set2") + 
  labs (
    x = "decade",
    y = "Surprise",
    fill = "Hypothesis"
  ) + 
  # scale_y_log10() +
  NULL

ggsave("ws-suprises-aggregate.pdf", width = 5, height = 5)
```
