---
title: "Thetas over time"
output:
    html_document:
    df_print: paged
theme: sandstone
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(viridis)
library(broom)
library(hrbrthemes)
theme_set(theme_ipsum(axis_title_size = 12))

knitr::opts_chunk$set(tidy = FALSE,
                      fig.width = 6,
                      fig.height = 5)

# library(thematic)
```

## Thetas

```{r read}
am_t_raw = read_csv(here::here("theta-csvs/theta_table_allmusic.csv.gz"))
am_names = read_csv(here::here("centrality-csvs/allmusic_names.csv"), col_names = c("id", "name"), col_types = "c") 
degrees = read_csv(here::here("graph-csvs/allmusic-degrees.csv"))
```

```{r}
glimpse(am_t_raw)
```

```{r}
skimr::skim(am_t_raw)
```

### Transform

```{r}
end_ranks = am_t_raw %>% 
  filter(ano == max(ano)) %>% 
  group_by(metric) %>% 
  select(artist, metric, end_theta_i = theta) %>% 
  ungroup()

am_t = am_t_raw %>% 
  filter(ano >= 1960) %>% 
  left_join(end_ranks, by = c("metric", "artist")) 

am_t = am_t %>% 
  left_join(am_names, by = c("artist" = "id")) %>% 
  left_join(degrees, by = c("artist", "ano" = "year")) %>% 
  filter(out_degree >= 5, in_degree >= 5)

am_t_wide = am_t %>% 
  select(artist, name, ano, theta, metric) %>% 
  pivot_wider(names_from = metric, values_from = theta)

with_history = am_t %>% 
  count(metric, artist) %>%
  filter(n > 2) %>% 
  distinct(metric, artist) %>% 
  count(artist) %>% 
  filter(n == 2) %>% 
  pull(artist)

am_t_h = am_t %>% 
  filter(artist %in% with_history)

am_t_h_wide = am_t_wide %>% 
  filter(artist %in% with_history)
```

### Explore

Some exemplary artists

```{r}
few_artists = function(d, n){
  d %>% 
    filter(metric == "disruption") %>% 
    sample_n(n) %>% 
    pull(artist)
}

examples_high_t = am_t_h %>% 
  group_by(metric, artist, name) %>%
  summarise(avg = mean(theta), count = n()) %>%
  filter(count > 2) %>%
  ungroup %>% 
  group_by(metric) %>% 
  top_n(3, avg) 
examples_high_t

examples_sd_t = am_t_h %>% 
  group_by(metric, artist, name) %>%
  summarise(sd = sd(theta)) %>%
  ungroup %>% 
  group_by(metric) %>% 
  top_n(3, sd) 


model_t = function(d){
  lm(theta ~ ano, data = d)
}

models = am_t_h %>%
  group_by(artist, metric) %>% 
  nest() %>% 
  mutate(model = map(data, model_t))

coefs = models %>% 
  mutate(coefs = map(model, tidy)) %>% 
  unnest(coefs)  %>% 
  filter(term == "ano")

examples_grow_disrupt = coefs %>% 
  select(artist, metric, estimate) %>% 
  filter(metric == "disruption") %>% 
  ungroup %>% 
  top_n(4, estimate)

examples_grow_disrupt
```

Everything over time

```{r}
plot_traj_sbs = function(d, artists = c(), names = c()){
  d %>% 
    filter(artist %in% artists | name %in% names) %>%
    ggplot(aes(x = ano, y = theta, color = name, group = artist)) +
    facet_wrap(~ metric) + 
    geom_line() + 
    geom_point()
}

am_t %>% 
  plot_traj_sbs(few_artists(., 5))

am_t %>% 
  plot_traj_sbs(examples_grow_disrupt$artist)

am_t %>% 
  plot_traj_sbs(names = c("Kraftwerk", "John Cage", "Sun Ra"))

am_t %>% 
  plot_traj_sbs(names = examples_sd_t$name) 
```

All thetas over time together

```{r}
am_t %>% 
  filter(metric == "pagerank") %>% 
  ggplot(aes(x = ano, y = end_theta_i, fill = theta)) + 
  facet_grid(. ~ metric) +
  geom_tile() + 
  scale_fill_viridis(direction =  -1, option = "inferno")

am_t %>% 
  filter(metric == "disruption") %>% 
  ggplot(aes(x = ano, y = end_theta_i, fill = theta)) + 
  facet_grid(. ~ metric) +
  geom_tile() + 
  scale_fill_viridis(direction =  -1, option = "inferno")
```

#### Trajectories on a xy plane.

```{r}
am_t_wide %>% 
  # filter(artist %in% few_artists(., 3)) %>%
  # filter(name %in% c("Kraftwerk")) %>%
  # filter(artist %in% examples_grow_disrupt$artist) %>%
  filter(name %in% examples_sd_t$name) %>%
  ggplot(aes(x = pagerank, y = disruption, color = name, group = artist)) +
  geom_line() + 
  geom_point(aes(size = ano, alpha = ano)) 
```

Are there correlations?

```{r}
am_t_wide %>% 
  filter(!is.na(pagerank), !is.na(disruption)) %>% 
  ggplot(aes(x = pagerank, y = disruption)) +
  geom_point(alpha = .2, size = .5) 
```

#### Some interesting trajectories

```{r}
am_t_wide %>% 
  filter(name %in% c("Stevie Wonder", "Jorge Ben", "Shirley Bassey", "Fela Kuti")) %>% 
  ggplot(aes(x = pagerank, y = disruption, group = artist)) +
  facet_wrap(~ name) +
  geom_line() + 
  geom_point(aes(size = ano, alpha = ano)) 
```

---

## Surprises

```{r}
am_s = read_csv(here::here("surprises-csvs/allmusic_surprises.csv.gz"))

am_s = am_s %>%
  mutate(
    surprise_disrupt = disrupt_surprise_hyp_growth_t +
      disrupt_surprise_hyp_past_t,
    surprise_pr = pagerank_surprise_hyp_growth_t + 
      pagerank_surprise_hyp_past_t,
    surprise_total = surprise_disrupt + surprise_pr
  ) %>% 
  filter(surprise_total != 0)
```

```{r}
glimpse(am_s)
```

Final dataframe

```{r}
allmusic =
  am_t_h_wide %>%
  filter(!is.na(pagerank + disruption)) %>% 
  left_join(
    select(am_s, name, year, matches("surprise")),
    by = c("ano" = "year", "name")
  ) %>% 
  filter(!is.na(surprise_total)) %>% 
  group_by(artist) %>% 
  arrange(ano) %>% 
  mutate(year_index = 1:n()) %>% 
  ungroup()

glimpse(allmusic)
```

### Agregate views

Surprise x year index

```{r}
allmusic %>% 
  group_by(artist) %>% 
  mutate(i = 1:n()) %>% 
  pivot_longer(matches("surprise"), names_to = "surprise") %>% 
  ggplot(aes(x = i, y = value, group = i)) +
  facet_wrap(~surprise, ncol = 2) +
  ggbeeswarm::geom_quasirandom(size = .3) +
  # geom_boxplot(coef = 1000) +
  scale_y_log10() +
  labs(x = "Year index from appearance onwards")
```



```{r}
allmusic %>% 
  group_by(artist) %>% 
  mutate(i = 1:n()) %>% 
  ggplot(aes(x = i, y = surprise_total)) +
  ggbeeswarm::geom_quasirandom(size = .5, alpha = .3) + 
  scale_y_log10() + 
  labs(x = "Year index from appearance onwards")

```

```{r}
allmusic %>% 
  group_by(artist) %>% 
  arrange(ano) %>% 
  mutate(i = 1:n()) %>% 
  mutate(delta = surprise_total - lag(surprise_total)) %>%  
  ggplot(aes(x = i, y = delta, group = i)) +
  # ggbeeswarm::geom_quasirandom(size = .3) +
  geom_boxplot(coef = 1e9) + 
  # geom_boxplot(coef = 1000) +
  labs(x = "Year index from appearance onwards") + 
  ylim(-10, 10)
```


```{r}
allmusic %>% 
  ggplot(aes(x = surprise_pr, y = surprise_disrupt)) + 
  geom_point(alpha = .25, size = .5) + 
  scale_x_log10(limits = c(1e-5, NA)) + 
  scale_y_log10(limits = c(1e-5, NA))

allmusic %>% 
  ggplot(aes(x = disrupt_surprise_hyp_past_t, y = disrupt_surprise_hyp_growth_t)) + 
  geom_point(alpha = .25, size = .5) + 
  scale_x_log10() + 
  scale_y_log10()


allmusic %>% 
  ggplot(aes(x = pagerank_surprise_hyp_past_t, y = pagerank_surprise_hyp_growth_t)) + 
  geom_point(alpha = .25, size = .5) + 
  scale_x_log10(limits = c(1e-6, 1e3)) + 
  scale_y_log10(limits = c(1e-6, 1e3))
```


#### highest surprise at some point


```{r}
plot_trajectory = function(artists, df){
  df %>% 
    filter(artist %in% artists) %>% 
    ggplot(aes(
      x = pagerank,
      y = disruption,
      color = year_index,
      fill = year_index,
      group = artist
    )) +
    facet_wrap( ~ name) +
    geom_line(size = 0.4, color = "black") +
    geom_point(aes(size = surprise_total), alpha = .8, shape = 21, color = "black") +
    #scale_color_continuous(trans = "reverse") +
    scale_fill_viridis(option = "magma", direction = -1) +
    scale_size_area(trans = "sqrt", max_size = 10) + 
    labs(
      x = expression(1 - theta * " for PageRank"),
      y = expression(1 - theta * " for Disruption"),
      fill = "Decades active",
      size = "Surprise"
    )
}
```


```{r warning=FALSE}
allmusic %>% 
  group_by(artist) %>% 
  summarise(surprise_total = max(surprise_total)) %>% 
  top_n(6, surprise_total) %>% 
  distinct(artist) %>% 
  pull(artist) %>% 
  plot_trajectory(allmusic)
```


### Highest *after* first year


```{r}
allmusic %>% 
  group_by(artist) %>% 
  arrange(ano) %>% 
  mutate(delta = surprise_total - lag(surprise_total)) %>%  
  summarise(surprise_total = max(delta, na.rm = T)) %>% 
  top_n(9 , surprise_total) %>% 
  distinct(artist) %>% 
  pull(artist) %>% 
  plot_trajectory(allmusic)
```




```{r}
allmusic %>% 
  filter(name %in% c("Luny Tunes", "Frank Sinatra", "Miles Davis")) %>% 
  pull(artist) %>% 
  plot_trajectory(allmusic)
```


#### highest average total surprise

```{r}
allmusic %>% 
  group_by(artist) %>% 
  summarise(surprise_total = mean(surprise_total)) %>% 
  top_n(6, surprise_total) %>% 
  distinct(artist) %>% 
  pull(artist) %>% 
  plot_trajectory(allmusic) +
  xlim(.3, 1)
```


Some interesting artists

```{r warning=FALSE}
allmusic %>%
  filter(
    name %in% c(
      # "Stevie Wonder",
      "Nina Simone",
      # "Dua Lipa",
      # "Shirley Bassey",
      "Björk",
      "Fela Kuti",
      # "Miles Davis",
      "Caetano Veloso"
    ),
    ano >= 1960
  ) %>%
  pull(artist) %>% 
  plot_trajectory(allmusic) +
  theme(legend.position = "bottom",
        plot.margin=grid::unit(c(0,5,0,0), "mm"))

  # xlim(.3, 1) + 
  NULL

ggsave(here::here("am-trajectories.pdf"), width = 5, height = 5)
```

```{r}
am_s_wide = allmusic %>% 
  pivot_longer(cols = matches("surprise"), names_to = "surprise") 
  
# am_s_wide %>% 
#   ggplot(aes(x = value)) + 
#   geom_density() +
#   facet_grid(surprise ~ ano) + 
#   scale_x_log10(limits = c(1e-4, 1)) 

allmusic %>% 
  ggplot(aes(x = surprise_pr, y = surprise_disrupt)) + 
  geom_point(alpha = .2, size = .5, color = "brown") + 
  facet_wrap(~ano) +
  scale_x_log10(limits = c(1e-6, 1e1)) + 
  scale_y_log10() +
  NULL
```

```{r}
rename_surprises = function(d){
  d %>% 
    mutate(
    s_class = if_else(grepl("disrupt",surprise), "Disruption", "Pagerank"),
    surprise = case_when(
      surprise == "pagerank_surprise_hyp_past_t" ~ "Past Rank",			
      surprise == "pagerank_surprise_hyp_growth_t"  ~ "Reg. Growth ",				
      surprise == "disrupt_surprise_hyp_past_t" ~ "Past Rank",			
      surprise == "disrupt_surprise_hyp_growth_t"~ "Reg. Growth ",				
      surprise == "surprise_disrupt" ~ "Disruption, sum",			
      surprise == "surprise_pr" ~ "Pagerank, sum"
    )
  )
}

am_s_wide %>% 
  filter(grepl("surprise", surprise), surprise != "surprise_total") %>% 
  rename_surprises() %>% 
  filter(value != 0) %>% 
  ggplot(aes(x = ano, y = value, group = ano, fill = s_class)) + 
  geom_violin() +
  geom_boxplot(width = 2.5, coef = 1e6, fill = "white") + 
  facet_wrap(~ surprise, ncol = 3) + 
  scale_y_log10(
                # limits = c(1e-5, 1e1)
                ) +
  theme(legend.position = "None") +
  scale_fill_ft() + 
  labs (
    x = "decade",
    y = "Surprise"
  ) 
```

```{r}
# library(showtext)
# ## Loading Google fonts (https://fonts.google.com/)
# font_add_google("Gochi Hand", "gochi")
# font_add_google("Schoolbell", "bell")
# 
# ## Automatically use showtext to render text
# showtext_auto()
```


```{r}
am_s_wide %>% 
  filter(grepl("surprise", surprise), !str_starts(surprise, "surprise")) %>% 
  rename_surprises() %>% 
  group_by(s_class, surprise, ano) %>% 
  summarise(value = sum(value)) %>% 
  ggplot(aes(x = ano, y = value, fill = surprise)) + 
  geom_area(alpha = .75) +
  facet_wrap(~ s_class, ncol = 1) + 
  theme(legend.position = "bottom",
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        ) +
  scale_fill_brewer(palette = "Set2") + 
  labs (
    x = "decade",
    y = "Surprise",
    fill = "Hypothesis"
  ) + 
  NULL

ggsave("am-suprises-aggregate.pdf", width = 5, height = 5)
```

```{r}
am_s_wide %>% 
  filter(ano %in% c(1990, 2000), str_starts(surprise, "surprise"), surprise != "surprise_total") %>% 
  group_by(ano, surprise) %>% 
  top_n(5, wt = value) %>% View
```

